Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-04-29 WRITEUP/ms.tex

Such massive gas inflows trigger gravitational instabilities that lead to the formation of molecular 
%mm complexes [first use of this term; did you want "cloud complexes" instead of "complexes" for agreement with abstract?] 
   cloud complexes (MCCs) 
that are typically more massive ($M_{\rm cl}\simeq 10^9$\,\Msun) and extended ($\simeq$ sub-kpc) than those observed in nearby galaxies (e.g., \citealt{Gabor13a, Hopkins14a, Inoue16a}).
%
Some theoretical works argue that the migration of such giant massive clumps are largely responsible for the buildup of the bulges of massive galaxies at redshift \z$\sim$\,0 \citep[e.g.,][]{Ceverino10a}.

@@

The virial theorem for a distribution of unmagnetized gas can be written as \citep{McKee92a}
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm cl}$, radius $R$, and root-mean-square velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions. 
%mm Neglecting magnetic field and 
Defining $P_{\rm ext}$ as the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M_{\rm cl} \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM_{\rm cl}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the right-hand side represent the
%mm  kinetic, external, and gravity pressure terms. 
    kinetic energy, external pressure, and gravitational potential energy terms.

Motivated by Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, we assume equilibrium (i.e.,\ ${\ddot{\mathcal{I}}}=0$), define the cloud surface density as
$\Sigma$\eq$M/\pi R^2$, and rewrite the previous equation as 
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \frac{3}{5} \pi G \Sigma \right)\,,
\label{eqn:v0}
\end{equation}
which further reduces to 
\begin{equation}
\frac{\sigma^2}{R} = \frac{\pi}{5} G \Sigma\,,
\label{eqn:SVE}
\end{equation}
if the external pressure $P_{\rm ext}=0$. For this case (often referred to as simple virial equilibrium) from the balance between kinetic and gravity terms we can define the virial parameter as 
\begin{equation}
\alpha_{\rm vir} \equiv  \frac{5\sigma^2R}{GM_{\rm cl}} = \frac{5\sigma_{\rm gas}^2}{\pi G \Sigma R}\,,
\label{eqn:alpha}
\end{equation}

@@

In cases where stellar component plays an important role in the dynamics, the virial parameter becomes
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G} \frac{M_{\rm cl} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm cl} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}

\AP{shouldn t it be something like}
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G (M_{\rm cl} + M_{\star})} \frac{M_{\rm cl} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm cl} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}
\AP{? Additionally, I am still unsure if the stellar dispersion should give a contribution to the alpha parameter.}

@@

The onset of gravitational instability is 
%mm considered to be 
tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. 
%mm [rearranged to put derivation before result]
For axisymmetric modes, the dispersion relation for the growth of density perturbations in a rotating, turbulent disk of finite thickness $h$ is described by
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma_{\rm disk}^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{\altomega}\frac{d}{d\altomega}\left(\altomega^2\Omega\right)\,.
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
In Equation~\ref{eqn:3Ddisp} the terms on the right hand side are related to rotation, self-gravity and internal pressure, respectively. Heuristically, the instability can be understood by considering the scale at which gravitational potential overcomes the internal energy. Gravity dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, differential rotation in disk galaxies can stabilize perturbations that might otherwise collapse for $L > L_{\rm rot}$, where $L_{\rm rot}$ is set by $\kappa$. As a result, disks are unstable to gravitational collapse on scales between $L_J < L < L_{\rm rot}$.

@@

In our stability analysis, we account for the combined effect of gas and stars 
%mm 
     \citep[derived exactly by][]{Rafikov01}, 
and for the non-negligible disk thickness. This is done by adopting an approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modifies the $Q$ parameter for gas and stars by accounting for the vertical velocity dispersion
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the finite disk thickness reduces the gravity in the vertical direction, thereby making it easier for a system to
maintain 
stability, and thus lowering the critical Toomre $Q_{\rm crit}$ from $\simeq$\,1 to 
0.67 \citep{Goldreich65a}.
%
On the other hand, including the contribution of the stellar component promotes gravitational instability, and thus increases
$Q_{\rm crit}$, more-so if the stars have low velocity dispersion.
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

% ------------------
\section{Numerical Simulations}\label{sec:sim}
The simulations used in this work are described by \citealt{Pallottini17a, Pallottini17b} and are only briefly summarized here.
%
\ncode{Serra}\footnote{Greenhouse in Italian.} is a suite of cosmological zoom-in simulations performed using Eulerian hydrodynamics and adaptive mesh refinement (AMR) techniques to achieve high spatial resolution in regions of interest (e.g., regions of high density).
%
In particular, it uses a modified version of \ncode{ramses} \citep{Teyssier02a} as the AMR backend. The simulation used here covers a comoving box of 20\,Mpc $h$\pmOne in size. The simulation zooms in on a target halo of mass $M_{\rm DM}\simeq10^{11} \Msun$. The Lagrangian region of the halo (2.1\,Mpc $h$\pmOne) has a dark matter mass resolution of $\simeq 6\times 10^4 \Msun$, equivalent at initial density to a baryonic mass resolution of $1.2 \times 10^4 \Msun$. This region is spatially refined with a quasi-Lagrangian criterion based on a mass threshold, so that a cell is refined if its total (dark+baryionic) mass exceedes the mass resolution by a factor of eight. The finest refined cell allowed in the zoom in region has size $l_{\rm cell}\simeq$\,30\,pc (at $z = 6$), i.e. sizes are comparable to the sizes of local giant molecular clouds \citep[e.g.,][]{Sanders85a, Federrath13a, Goodman14a},
The models include a non-equilibrium chemical network following e$^{-}$, H$^+$, H$^-$, He, He$^+$, He$^{++}$, H$_2$, and H$_2^+$ \citep{Grassi14a,Bovino16a}. Of particular importance here is that abundances are calculated using an on-the-fly non-equilibrium formation of molecular hydrogen scheme described by \citet{Pallottini17a}.

Star formation is modeled using an H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar tracks from \ncode{starburst99} and include stellar feedback from supernovae (SNe) as well as winds from OB and asymptotic giant branch stars. Coupling to the gas is implemented via a sub-grid model for blastwaves, which accounts for radiative energy losses inside the cell. The remaining energy is injected into the ISM in kinetic and thermal energy forms. Radiation pressure on dust and gas is also included (see \citealt{Pallottini17a} for details). Photoionization from local sources is neglected in this simulation \citep[see][for its effect]{pallottini:2019,decataldo:2019}.

Note that in the simulation the gas acquire energy from feedback processes in both thermal and non-thermal form \citep{agertz:2013}, with the latter being a term that mimics unresolved turbulence; the non-thermal energy is not affected by cooling but rather is dissipated on the eddy turn-over time scale \citep{maclow1999turb}; see \citet{Pallottini17b} for further details.

@@

Given the stochastic nature of \flower in its star formation history, we mostly focus on
two of its most extreme evolutionary stages in this work (see \Sec{singless}).
These phases correspond to (a) an intensely accreting phase and
(b) a starburst phase (\Fig{SFH}). We are interested in determining whether 
MCC properties are sensitive to these different dynamical
conditions. For completeness, we also show the scaling relations examined for the other evolutionary stages 
traced in the simulation (see \Sec{ncut}). 
In \Sec{rot}, we investigate the importance of rotation support from large-scale motions in the MCC dynamics by comparing their properties 
in the disturbed phase of \flower and in the disk-like ordered phase.

@@

To identify the molecular complexes, we use a customized version of
the clump-finding algorithm available in the \ncode{python} package
\ncode{yt} \citep{Turk11a}, which was initially described in
\citet{Smith09a}, 
%mm but this function has been modified since then.
     and modified since.
The latest version of the default \ncode{yt} clump finder decomposes
the zones of the simulation into non-overlapping tiles, which are
stored in a 
%mm $k$-dimensional tree ($k$-D tree). [unused acronym]
     three-dimensional tree that can be processed using $k$-dimensional tree algorithms.
It then identifies the contours of a variable field (here, the density
field) within a tile and connects them across the tiles. In the
customized version used for this study, we
%mm modify the function to enhance 
     enhanced
the stability of the code.
%
Due to the nature of our AMR simulation, we regrid the simulation data into uniform grids. The grid size is defined based on the highest resolution of the simulation data, i.e., the less refined regions are supersampled in the resulting uniform grids.

In the clump-finding process (in position-position-position space; PPP), we employ a set of different density
thresholds defined based on the molecular hydrogen density of \flower
at different evolutionary stages ($z=6.0$--7.2).
%
We note that this process is 
the three-dimensional analog
to identifying molecular structures based on the noise levels 
     in position-position-velocity (PPV) maps that
observers obtain with telescopes, using molecular line tracers such as
CO, CS, and HCN.
%mm , as commonly adopted in observational studies (e.g., 
     This is commonly done by 
identifying clumps based on/after applying S/N-clipping, using tools such as \ncode{aips}'s task \ncode{serch}, \ncode{clumpfind}, and \ncode{cprops}; \citealt{Williams94a, Oka01a, Rosolowsky06a, Rosolowsky08a}).
%
%We note that, owing to the nature of \obs, such structures are identified in position-position-velocity (PPV) space, whereas in simulations, one has the full 6D spatial-kinematic information, and can therefore cleanly identify structures directly using the density field in position-position-position (PPP) space.
Existing studies find a good correspondence in the dynamical
properties extracted in PPV- versus PPP-space 
%mm
   for well-isolated structures 
(\citealt{Ballesteros-Paredes02a, Heitsch09a, Shetty10a, Beaumont13a, Pan15a}, but see
%mm
    \citealt{Ballesteros-Paredes02a} and 
\citealt{Shetty10a} for a discussion on caveats and limitations).
% large scale structure in PPP may be identified as numerous lower mass structures in the PPV cube due to gradients in the LOS and conversly, superposition when spatially distinct components with the same velocity are "merged" into one component in the PPV space when observed along the LOS (see Fig. 1 of Beaumount13a). (may be worth putting more thoughts on this after the report is due).

@@

In \Fig{h2density}, we show the distribution of H$_2$
number density $n_{\rm H2}$ for \flower~ during its accreting phase,
including the contribution from the gas within 3.5 kpc from the
galaxy center. 
%
We note that the distribution is almost flat for $n_{\rm
  H2}\gtrsim1$\,\cc and it samples the range of densities where clumps
are found based on morphological analysis\footnote{For $n_{\rm
    H2}\gtrsim1$\,\cc, the fourth Minkowski functional of the H$_{2}$
  density field is significantly larger than zero. This implies that
  the field is made of isolated components. See Fig. 6 in
  \citet{Pallottini17b}.} \citep{Pallottini17b}. We identify MCCs by
applying 10 equally-spaced volumetric H$_2$ density cuts of ${(n_{\rm
    cut}/{1 {\rm cm}^{-3})}}\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08,
6.81, 11.36, 18.96, 31.62]$ to each snapshot. 
We also tested variations in
  $n_{\rm cut}$ range and found no qualitative differences (see
  \Sec{ncut}).
%mm
\MM{I am confused how H$_2$ can survive at such low number
  densities.  Is there some sort of hidden clumping factor at work
  here? Or is this due to the neglect of local photodissociating
  radiation? Either way, the point should be mentioned in section 3, I
guess.}
\DL{ncut should be more like molecular gas density, and H2 fraction in each molecular gas is not 100\%, so the ncut here are lower than
 what one expects for purely H2 regions. No clumping factor.}

To visually display 
the clump finding procedure, we overplot the molecular structures identified using a subset of the H$_2$ density cuts ($n_{\rm cut}$\eq0.53, 5.81, and 18.96\,\cc) on the H$_2$ density maps (\Fig{MCC}).
Since the molecular structures identified could easily appear as overlapping structures depending on the viewing angle, we also plot them in different three-dimensional projections (right panels of \Fig{MCC}) so that one can more easily see that they are collections of disjoint structures.
%
We repeat this identification process for 14 evolutionary stages between redshift \z$\in$[6.0, 7.2], spaced by $\Delta t$\eq15\,Myr.

Note that we impose the additional constraint that an identified
structure must be composed of at least 10 cells. We caution that an
important caveat of such a constraint is that we can only examine the
parameter space of cloud complexes of 
radius $R\gtrsim 40$\,pc,
because of the resolution limit of the simulation.

@@

% Molecular clouds are the natal place for \SF, their structure and dynamics hold important clues to understanding
% the mechanisms and physics of formation and evolution of molecular structures and \SF.
Upon identifying the molecular structures, we extract properties such
as the gas mass $M_{\rm cl}$, effective size $R$, Mach number
$\mathcal{M}$, velocity dispersion $\sigma$
, and gas surface density $\Sigma_{\rm gas}$ to examine their dynamics.

The mass of an MCC is calculated from the uniformly-gridded 3D density field, integrating over the MCC volume $V$. The effective size is defined assuming spherical geometry, i.e. $R \equiv (3 V /4 \pi)^{1/3}$.
The full velocity dispersion of MCCs is calculated from 
the bulk velocity field $\sigma_{\rm bulk}$, non-thermal velocity dispersion $\sigma_{\rm NT}$, and thermal sound speed $c_s$, following
\begin{equation}
\sigma^2 = \sigma_{\rm NT}^2 + \sigma_{\rm bulk}^2 + c_s^2.
\label{eqn:veldisp}
\end{equation}
%
In \obs of MCCs, the linewidth contribution of dense gas 
exceeds that from the diffuse gas. We therefore calculate each term on
the right-hand side of Equation~\ref{eqn:veldisp} as a density-weighted quantity. 
For instance, the bulk velocity dispersion is calculated as follows
\begin{equation}
\sigma_{\rm bulk}^2 = \frac{1}{{3}}\frac{\sum_{i} \rho_i \left|\mathbf{v}_i - \mathbf{\bar v}\right|^2}{\sum_i \rho_i}\,,
\end{equation}
where the sum is done for the cells composing each MCC.
Similarly, the non-thermal velocity dispersion is calculated as follows
\begin{equation}
\sigma_{\rm NT}^2 = \frac{\sum_{i} \rho_i \left(P_{\rm NT, i}~k_B / M_p / {\bar \rho_i} \right)^2}{\sum_i \rho_i}\,,
\end{equation}
and the thermal sound speed is calculated through
\begin{equation}
%        x = (self.Pressure * self.k_B_erg / C.m_p.cgs.value) / self.avg_density
c_s^2 = \frac{\sum_{i} \rho_i \left(P_i~k_B / M_p / {\bar \rho_i} \right)^2}{\sum_i \rho_i}\,,   % so that cs^2 has dimension of [erg/g]
\end{equation}
where $P_{\rm NT, i}$ and $P_i$ are the non-thermal and thermal pressures of cell $i$ in units of [K\,\cc], respectively, $k_B$ is the Boltzmann constant, and 
$M_p$ is the mass of a proton.

>>>>

Such massive gas inflows trigger gravitational instabilities that lead to the formation of molecular 
%mm complexes [first use of this term; did you want "cloud complexes" instead of "complexes" for agreement with abstract?] 
   cloud complexes (MCCs) 
that are typically more massive ($M_{\rm gas}\simeq 10^9$\,\Msun) and extended ($\simeq$ sub-kpc) than those observed in nearby galaxies (e.g., \citealt{Gabor13a, Hopkins14a, Inoue16a}).
%
Some theoretical works argue that the migration of such giant massive clumps are largely responsible for the buildup of the bulges of massive galaxies at redshift \z$\sim$\,0 \citep[e.g.,][]{Ceverino10a}.

@@

The virial theorem for a distribution of unmagnetized gas can be written as \citep{McKee92a}
\begin{equation}\label{eqn:virial_th_general}
\frac{1}{2}\ddot{\mathcal{I}} = 2(\mathcal{T} - \mathcal{T}_{\rm ext}) + \mathcal{W},
\end{equation}
where $\ddot{\mathcal{I}}$ is the second time derivative of the Lagrangian moment of inertia, $\mathcal{T}$ is the internal energy of the gas, $\mathcal{T}_{\rm ext}$ is external pressure support, and $\mathcal{W}$ is the gravitational energy.
%
Let us specialize to the case of a spherical self-gravitating molecular cloud of mass $M_{\rm gas}$, radius $R$, and root-mean-square velocity dispersion $\sigma$, accounting for both thermal and turbulent contributions. 
%mm Neglecting magnetic field and 
Defining $P_{\rm ext}$ as the external pressure, \Eq{virial_th_general} can be written as
\begin{equation}
\frac{1}{2}{\ddot{\mathcal{I}}} = 3 M_{\rm gas} \sigma^2 - 4\pi P_{\rm ext} R^3 - \Gamma\frac{GM_{\rm gas}^2}{R}\,,
\label{eqn:virial}
\end{equation}
where $\Gamma$ is a geometrical factor that is equal to 3/5 for a uniform sphere; in \Eq{virial} the terms on the right-hand side represent the
%mm  kinetic, external, and gravity pressure terms. 
    kinetic energy, external pressure, and gravitational potential energy terms.

Motivated by Larson's linewidth-size relation \citep{Larson81a} and the work by \citet{Heyer09a}, we assume equilibrium (i.e.,\ ${\ddot{\mathcal{I}}}=0$), define the cloud surface density as
$\Sigma$\eq$M/\pi R^2$, and rewrite the previous equation as 
\begin{equation}
\frac{\sigma^2}{R} = \frac{1}{3}\left(\frac{4P_{\rm ext}}{\Sigma} + \frac{3}{5} \pi G \Sigma \right)\,,
\label{eqn:v0}
\end{equation}
which further reduces to 
\begin{equation}
\frac{\sigma^2}{R} = \frac{\pi}{5} G \Sigma\,,
\label{eqn:SVE}
\end{equation}
if the external pressure $P_{\rm ext}=0$. For this case (often referred to as simple virial equilibrium) from the balance between kinetic and gravity terms we can define the virial parameter as 
\begin{equation}
\alpha_{\rm vir} \equiv  \frac{5\sigma^2R}{GM_{\rm gas}} = \frac{5\sigma_{\rm gas}^2}{\pi G \Sigma R}\,,
\label{eqn:alpha}
\end{equation}

@@

In cases where stellar component plays an important role in the dynamics, the virial parameter becomes
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}

\AP{shouldn t it be something like}
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R}{G (M_{\rm gas} + M_{\star})} \frac{M_{\rm gas} \sigma_{\rm gas}^2 + M_{\star} \sigma_\star^2}
					       {M_{\rm gas} + M_{\star} }\,,
\label{eqn:alpha_tot}
\end{equation}
\AP{? Additionally, I am still unsure if the stellar dispersion should give a contribution to the alpha parameter, naively I would write it as}
\begin{equation}
\alpha_{\rm vir, tot} \equiv \frac{5 R \sigma_{\rm gas}^2}{G (M_{\rm gas} + M_{\star})}
\label{eqn:alpha_tot}
\end{equation}

@@

The onset of gravitational instability is tightly connected to \SF \citep[e.g.,][]{Kennicutt89a, Wang94a, Li05b, Li06a}. For axisymmetric modes, the dispersion relation for the growth of density perturbations in a rotating, turbulent disk of finite thickness $h$ is described by \AP{there might be too many $\sigma$}
\begin{equation}
\omega^2 = \kappa^2 - \frac{2\pi G \Sigma |k|}{1 + |k| h} + \sigma_{\rm disk}^2 k^2\,,
\label{eqn:3Ddisp}
\end{equation}
where $k$ is the wavenumber and $\kappa$ is the epicyclic frequency, defined as:
\begin{equation}
\kappa^2\equiv\frac{2\Omega}{\altomega}\frac{d}{d\altomega}\left(\altomega^2\Omega\right)\,.
\label{eqn:kappa}
\end{equation}
\citep{Romeo92a}.
In Equation~\ref{eqn:3Ddisp} the terms on the right hand side are related to rotation, self-gravity and internal pressure, respectively. Heuristically, the instability can be understood by considering the scale at which gravitational potential overcomes the internal energy. Gravity dominates at scales $L > L_J$, where $L_J$ is the Jeans length. However, differential rotation in disk galaxies can stabilize perturbations that might otherwise collapse for $L > L_{\rm rot}$, where $L_{\rm rot}$ is set by $\kappa$. As a result, disks are unstable to gravitational collapse on scales between $L_J < L < L_{\rm rot}$.

@@

In our stability analysis, we account for the combined effect of gas and stars \citep[derived exactly by][]{Rafikov01}, and for the non-negligible disk thickness. This is done by adopting an approximation for an effective two-component $Q_{\rm eff}$ parameter \citep[i.e.,][see also \citealt{Inoue16a}]{Romeo11a}.
%
The effect of disk thickness modifies the $Q$ parameter for gas and stars by accounting for the vertical velocity dispersion
\begin{equation}
T_{x} = \left\{
		\begin{array}{lccr}
			{\displaystyle 0.8 + 0.7\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}      && & \mbox{if\ } \sigma_z \lesssim 0.5 \times \sigma_r \\ [1.25em]
			{\displaystyle 1 + 0.6\left(\frac{\sigma_{z}}{\sigma_{r}}\right)}        & & & \mbox{if\ } \sigma_z \gtrsim 0.5 \times \sigma_r
\\
		\end{array}
	\right.
\end{equation}
and
\begin{equation}
Q^{\rm thick}_{x} = T_{x} Q\,,
\end{equation}
with $x$ indicating either gas or stars.
% Note that w/ thickness, there's the a self-gravity dilution term in the dispersion relation, which in effect reduceds the surface gravity in the vertical direction. As such, it is easier to be stable against gravitaional instability. Thus, the criterion for instability $Q$ is reduced.
%
The combined effect of gas and stars can then be accounted for by writing
\begin{equation}\label{eqn:q_eff}
Q^{-1}_{\rm eff} =  \left\{
				\begin{array}{lccr}
					     {\displaystyle\frac{w}{Q^{\rm thick}_{\star}} + \frac{1}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ }  Q^{\rm thick}_{\star} \geq Q^{\rm thick}_{\rm gas} \\ [0.75em]
                                               {\displaystyle\frac{1}{Q^{\rm thick}_{\star}} + \frac{w}{Q^{\rm thick}_{\rm gas}}}      & & & \mbox{if\ } Q^{\rm thick}_{\star} \leq Q^{\rm thick}_{\rm gas} \\
				\end{array}
			    \right.
\end{equation}
where the relative weight $w$ is defined as
\begin{equation}
w\equiv\frac{2 \sigma_{\star} \sigma_{\rm gas}}{\sigma_{\star}^2 + \sigma_{\rm gas}^2},
\end{equation}
Conceptually, the finite disk thickness reduces the gravity in the vertical direction, thereby making it easier for a system to maintain stability, and thus lowering the critical Toomre $Q_{\rm crit}$ from $\simeq$\,1 to 0.67 \citep{Goldreich65a}.
%
On the other hand, including the contribution of the stellar component promotes gravitational instability, and thus increases
$Q_{\rm crit}$, more-so if the stars have low velocity dispersion.
As a rule of thumb, $Q_{\rm crit}\eq1.34$ for $Q_{\rm gas}$\eq$Q_\star$.  % 0.67 * 2

% ------------------
\section{Numerical Simulations}\label{sec:sim}
The simulations used in this work are described by \citealt{Pallottini17a, Pallottini17b} and are briefly summarized here.
%
\ncode{Serra}\footnote{Greenhouse in Italian.} is a suite of cosmological zoom-in simulations performed using Eulerian hydrodynamics and adaptive mesh refinement (AMR) techniques to achieve high spatial resolution in regions of interest (e.g., regions of high density).
%
In particular, it uses a modified version of \ncode{ramses} \citep{Teyssier02a} as the AMR backend. The simulation used here covers a comoving box of 20\,Mpc $h$\pmOne in size. The simulation zooms in on a target halo of mass $M_{\rm DM}\simeq10^{11} \Msun$. The Lagrangian region of the halo (2.1\,Mpc $h$\pmOne) has a dark matter mass resolution of $\simeq 6\times 10^4 \Msun$, equivalent at initial density to a baryonic mass resolution of $1.2 \times 10^4 \Msun$. This region is spatially refined with a quasi-Lagrangian criterion based on a mass threshold, so that a cell is refined if its total (dark+baryionic) mass exceedes the mass resolution by a factor of eight. The finest refined cell allowed in the zoom in region has size $l_{\rm cell}\simeq$\,30\,pc (at $z = 6$), i.e. sizes are comparable to the sizes of local giant molecular clouds \citep[e.g.,][]{Sanders85a, Federrath13a, Goodman14a},
The models include a non-equilibrium chemical network following e$^{-}$, H$^+$, H$^-$, He, He$^+$, He$^{++}$, H$_2$, and H$_2^+$ \citep{Grassi14a,Bovino16a}. Of particular importance here is that abundances are calculated using an on-the-fly non-equilibrium formation of molecular hydrogen scheme described by \citet{Pallottini17a}.

Star formation is modeled using an H$_2$-based prescription of the Schmidt-Kennicutt relation \citep{Krumholz09a}. We adopt stellar tracks from \ncode{starburst99} and include stellar feedback from supernovae (SNe) as well as winds from OB and asymptotic giant branch stars. Coupling to the gas is implemented via a sub-grid model for blastwaves, which accounts for radiative energy losses inside the cell. The remaining energy is injected into the ISM in kinetic and thermal energy forms. Radiation pressure on dust and gas is also included. Photoionization from local sources is neglected in this simulation \citep[see][for its effect]{pallottini:2019,decataldo:2019}.
%
Via feedback processes the gas can acquire energy both thermal and non-thermal form \citep{agertz:2013}; the latter is a term that mimics unresolved turbulence: such non-thermal energy is not affected by cooling but rather is dissipated on the eddy turn-over time scale \citep{maclow1999turb}. See \citet{Pallottini17b} for further details.

@@

Given the stochastic nature of \flower in its star formation history, we mostly focus on few of its most extreme evolutionary stages in this work (see \Sec{singless}).
%
These phases correspond to (a) an intensely accreting phase and (b) a starburst phase (\Fig{SFH}). We are interested in determining whether MCC properties are sensitive to these different dynamical conditions. For completeness, we also show the scaling relations examined for the other evolutionary stages traced in the simulation (see \Sec{ncut}).

In \Sec{rot}, we investigate the importance of rotation support from large-scale motions in the MCC dynamics by comparing their properties 
in the disturbed phase of \flower and in the disk-like ordered phase.

@@

To identify the molecular complexes, we use a customized version of the clump-finding algorithm available in the \ncode{python} package \ncode{yt} \citep{Turk11a}, which was initially described in \citet{Smith09a} and modified since.
%
The latest version of the default \ncode{yt} clump finder decomposes the zones of the simulation into non-overlapping tiles, which are stored in a three-dimensional tree that can be processed using $k$-dimensional tree algorithms. It then identifies the contours of a variable field (here, the density field) within a tile and connects them across the tiles. In the customized version used for this study, we enhanced the stability of the code.
%
Due to the nature of our AMR simulation, we regrid the simulation data into uniform grids. The grid size is defined based on the highest resolution of the simulation data, i.e., the less refined regions are supersampled in the resulting uniform grids.

In the clump-finding process (in position-position-position space; PPP), we employ a set of different density thresholds defined based on the molecular hydrogen density of \flower at different evolutionary stages ($z=6.0$--7.2).
%
We note that this process is the three-dimensional analog to identifying molecular structures based on the noise levels in position-position-velocity (PPV) maps that observers obtain with telescopes, using molecular line tracers such as CO, CS, and HCN. This is commonly done by 
identifying clumps based on/after applying S/N-clipping, using tools such as \ncode{aips}'s task \ncode{serch}, \ncode{clumpfind}, and \ncode{cprops}; \citealt{Williams94a, Oka01a, Rosolowsky06a, Rosolowsky08a}).
%
%We note that, owing to the nature of \obs, such structures are identified in position-position-velocity (PPV) space, whereas in simulations, one has the full 6D spatial-kinematic information, and can therefore cleanly identify structures directly using the density field in position-position-position (PPP) space.
Existing studies find a good correspondence in the dynamical properties extracted in PPV- versus PPP-space for well-isolated structures (\citealt{Ballesteros-Paredes02a, Heitsch09a, Shetty10a, Beaumont13a, Pan15a}, but see \citealt{Ballesteros-Paredes02a} and \citealt{Shetty10a} for a discussion on caveats and limitations).
% large scale structure in PPP may be identified as numerous lower mass structures in the PPV cube due to gradients in the LOS and conversly, superposition when spatially distinct components with the same velocity are "merged" into one component in the PPV space when observed along the LOS (see Fig. 1 of Beaumount13a). (may be worth putting more thoughts on this after the report is due).

@@

In \Fig{h2density}, we show the distribution of H$_2$ number density $n_{\rm H2}$ for \flower~ during its accreting phase, including the contribution from the gas within 3.5 kpc from the galaxy center. 
%
We note that the distribution is almost flat for $n_{\rm H2}\gtrsim1$\,\cc and it samples the range of densities where clumps are found based on morphological analysis\footnote{For $n_{\rm H2}\gtrsim1$\,\cc, the fourth Minkowski functional of the H$_{2}$ density field is significantly larger than zero. This implies that the field is made of isolated components. See Fig. 6 in \citet{Pallottini17b}.} \citep{Pallottini17b}.
%
In each evolutionary stage, we identify MCCs by applying density cuts to the H$_2$ density distriution, i.e. $n_{\rm H2}\geq n_{\rm cut}$. We select 10 equally-spaced cuts in log scale: ${(n_{\rm cut}/{1 {\rm cm}^{-3})}}\eq[0.32, 0.53, 0.88, 1.45, 2.45, 4.08, 6.81, 11.36, 18.96, 31.62]$. We also tested variations in $n_{\rm cut}$ range and found no qualitative differences (see \Sec{ncut}). Note that with these choice we are including in the MCC also material that is not fully molecular ($n = 0.5 n_{\rm H2}$).

To visually display the clump finding procedure, we overplot the molecular structures identified using a subset of the H$_2$ density cuts ($n_{\rm cut}$\eq0.53, 5.81, and 18.96\,\cc) on the H$_2$ density maps (\Fig{MCC}). Since the molecular structures are identified in the 3D H2 density field, they can appear as overlapping structures depending on the viewing angle; thus we also plot them in different projections (right panels of \Fig{MCC}) so that the identification can be more easily appreciated.
%
We repeat this identification process for 14 evolutionary stages between redshift \z$\in$[6.0, 7.2], spaced by $\Delta t$\eq15\,Myr.

We impose the additional constraint that an identified structure must be composed of at least 10 cells. We caution that an important caveat of such a constraint is that we can only examine the parameter space of cloud complexes of radius $R\gtrsim 40$\,pc, because of the resolution limit of the simulation.

@@

Upon identifying the molecular structures, we extract properties such as the gas mass $M_{\rm gas}$, effective size $R$, Mach number
$\mathcal{M}$, velocity dispersion $\sigma_{\rm gas}$, and gas surface density $\Sigma_{\rm gas}$ to examine their dynamics.
